---
title: "IP_W13_Part 2"
author: "Geoffrey Chege"
date: '2022-03-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Problem Definition

Kira Plastinina is a Russian brand that is sold through a defunct chain of retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia. The brand’s Sales and Marketing team would like to understand their customer’s behavior from data that they have collected over the past year. More specifically, they would like to learn the characteristics of customer groups.

# 2. Steps Taken

- Problem Definition
- Data Sourcing
- Check the Data
- Perform Data Cleaning
- Perform Exploratory Data Analysis:
  - Univariate
  - Bivariate
  - Multivariate
- Implement the Solution
- Challenge the Solution
- Follow up Questions

# 3. Data Sourcing

The dataset for this Independent project can be found here [http://bit.ly/EcommerceCustomersDataset 

- The dataset consists of 10 numerical and 8 categorical attributes. The 'Revenue' attribute can be used as the class label.
- "Administrative", "Administrative Duration", "Informational", "Informational Duration", "Product Related" and "Product Related Duration" represents the number of different types of pages visited by the visitor in that session and total time spent in each of these page categories. The values of these features are derived from the URL information of the pages visited by the user and updated in real-time when a user takes an action, e.g. moving from one page to another. 
- The "Bounce Rate", "Exit Rate" and "Page Value" features represent the metrics measured by "Google Analytics" for each page in the e-commerce site. 
- The value of the "Bounce Rate" feature for a web page refers to the percentage of visitors who enter the site from that page and then leave ("bounce") without triggering any other requests to the analytics server during that session. 
- The value of the "Exit Rate" feature for a specific web page is calculated as for all pageviews to the page, the percentage that was the last in the session.
- The "Page Value" feature represents the average value for a web page that a user visited before completing an e-commerce transaction. 
- The "Special Day" feature indicates the closeness of the site visiting time to a specific special day (e.g. Mother’s Day, Valentine's Day) in which the sessions are more likely to be finalized with the transaction. The value of this attribute is determined by considering the dynamics of e-commerce such as the duration between the order date and delivery date. For example, for Valentina’s day, this value takes a nonzero value between February 2 and February 12, zero before and after this date unless it is close to another special day, and its maximum value of 1 on February 8. 
- The dataset also includes the operating system, browser, region, traffic type, visitor type as returning or new visitor, a Boolean value indicating whether the date of the visit is weekend, and month of the year.

## 4. Installing and loading Necessary Packages
```{r eval=FALSE, include=FALSE}

install.packages("DataExplorer")
install.packages("Hmisc")
install.packages("pastecs")
install.packages("iterators")
install.packages("caret")
install.packages("caretEnsemble")
install.packages("ggplot2")
install.packages("e1071")
install.packages("factoextra")
install.packages("randomForest")
install.packages("ggcorrplot")
install.packages('ranger')
install.packages("Rtsne")
install.packages('caTools')
install.packages('rpart')
install.packages('rpart.plot')
install.packages('psych')
install.packages("devtools")
```

```{r include=FALSE}

library(tidyr)
library(magrittr)
library(warn = -1)
library(RColorBrewer)
library(ggplot2)
library(lattice)
library(corrplot)
library(DataExplorer)
library(Hmisc)
library(pastecs)
library(psych)
library(factoextra)
library(Rtsne)
library(caret)
library("rpart.plot")

```

## 5. Check the Data

```{r df}

df <- read.csv("C:/Users/user/Downloads/online_shoppers_intention.csv")
head(df)

```


## 6. Data Cleaning

### 6.1 Missing Values

```{r}

# checking for missing values

colSums(is.na(df))

```

```{r}

# dropping null values

df <- na.omit(df)

```

```{r}

# confirming there are no null values

colSums(is.na(df))

```

### 6.2 Checking for duplicates

```{r duplicates}
duplicates <- df[duplicated(df),]
duplicates
```

I will drop the duplicates

```{r}
# eliminating for duplicates
df <- df[!duplicated(df), ]
```


```{r}
### Dataset structure
str(df)
```

### 6.3 Changing columns to factors

```{r}
# changing columns to factors

df$Month <- factor(df$Month)
df$VisitorType <- factor(df$VisitorType)
df$Weekend <- factor(df$Weekend)
df$Revenue <- factor(df$Revenue)

```

```{r}
### Dataset structure
str(df)
```

## 7. Exploratory Data Analysis

#### A function to determine the mode

```{r mode}
mode <- function(v){
  uniq <- unique(v)
  uniq[which.max(tabulate(match(v,uniq)))]
}
```

#### Summary statistics of the columns

```{r}

summary(df)

```

####  Description of Columns

```{r}

describe(df)

```

### Univariate Analysis

#### Administrative Column

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 2.34
  - Median: 1
  - Skewness: 1.95
  - Kurtosis: 4.63

- The mode is:

```{r}
mode(df$Administrative)
```

#### Informational Column

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 0.51
  - Median: 0
  - Skewness: 4.01
  - Kurtosis: 26.64

- The mode is:

```{r}
mode(df$Informational)
```

#### ProductRelated Column

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 32.06
  - Median: 18
  - Skewness: 4.33
  - Kurtosis: 31.04

- The mode is:

```{r}
mode(df$ProductRelated)
```

#### BounceRates

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 0.02
  - Median: 0.00
  - Skewness: 3.15
  - Kurtosis: 9.25

- The mode is:

```{r}
mode(df$BounceRates)
```

#### ExitRates

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 0.04
  - Median: 0.03
  - Skewness: 2.23
  - Kurtosis: 4.62

- The mode is:

```{r}
mode(df$ExitRates)
```

#### PageValues

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 5.95
  - Median: 0
  - Skewness: 6.35
  - Kurtosis: 64.93
  
- The mode is:

```{r}
mode(df$PageValues)
```

#### SpecialDay

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 0.06
  - Median: 0
  - Skewness: 3.28
  - Kurtosis: 9.78
  
- The mode is:

```{r}
mode(df$SpecialDay)
```

#### Month

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 6.17
  - Median: 7
  - Skewness: -0.83
  - Kurtosis: -0.37
  
- The mode is:

```{r}
mode(df$Month)
```

#### OperatingSystems

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 2.12
  - Median: 2
  - Skewness: 2.03
  - Kurtosis: 10.27
  
- The mode is:

```{r}
mode(df$OperatingSystems)
```

#### Browser

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 2.36
  - Median: 2
  - Skewness: 3.22
  - Kurtosis: 12.53
  
- The mode is:

```{r}
mode(df$Browser)
```

#### Region

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 3.15
  - Median: 3
  - Skewness: 0.98
  - Kurtosis: -0.16
  
- The mode is:

```{r}
mode(df$Region)
```

#### TrafficType

- From the summary and description, we can gather the following about the administrative column:
  - Mean: 4.07
  - Median: 2
  - Skewness: 1.96
  - Kurtosis: 3.47
  
- The mode is:

```{r}
mode(df$TrafficType)
```

#### Distributions

```{r}
plot_histogram(df)
```

```{r}
plot_bar(df)
```

### Bivariate Analysis

All the data profiling statistics will be organized into the report below

```{r}
create_report(df)
```

The link for the report is here: "file:///C:/Users/user/Documents/IP_W13_Part%202/report.html#qq-plot"

## 8. Implementing the Solution

### K-Means Clustering

Step 1: One hot encoding of the factor variables.

```{r}

# # One hot encoding of the factor variables.

dmy = dummyVars(" ~ .", data = df)

df2 = data.frame(predict(dmy, newdata = df))

```


```{r}
# Checking the data types of each attribute
sapply(df2, class)
```

Step 2: We are instructed to use Revenue as the class label, Hence we will remove it and store it in another variable

```{r}
# Step 2
# We are instructed to use Revenue as the class label,
# Hence we will remove it and store it in another variable

df2_copy <- df2[, -c(30:31)]
df.class<- df[, "Revenue"]

df2_copy_copy <- df2[, -c(30,31)]

```


```{r}
# Previewing the copy dataset with dummies
head(df2_copy)
```

Step 3: Determining whether to Normalize or Scale the data


#### Scaling:

```{r}
# This is important to ensure that no particular attribute, has more impact on clustering algorithm than others

df2_scaled <- scale(df2_copy)
```


```{r}
# After scaling the data lets see what we find in the output
summary(df2_scaled)
```

It is evident that there are some attributes still with large values compared to others.
Scaling makes the data changes the data to have a mean 0.
We will normalize the data and see if we get different results.



#### Normalizing:

```{r}
# Normalizing the a copy of the original data

df2_norm <- as.data.frame(apply(df2_copy, 2, function(x) (x - min(x))/(max(x)-min(x))))
```


```{r}
# summary of the normalized data.
summary(df2_norm)
```

Here, we have a maximum value of 1 and minimum value of 0s and mean of close to zero in all attributes.
We will use the NORMALIZED dataset for clustering.


```{r}
# Searching for the optimal number of clusters
# # Elbow method
fviz_nbclust(df2_norm, kmeans, method = "wss") +
    geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")

```

```{r}
# Applying K-Means  Clustering algorithm 
# Using 3 centroids as K=3

result <- kmeans(df2_norm, 10)
```

```{r}
# Previewing the number of records in each cluster

result$size
```

```{r}
# Viewing the cluster center datapoints by each attribute

result$centers
```

```{r}
# Plotting two variables to see how their data points 
# have been distributed in the cluster
# Product Related, vs Product Related Duration

plot(df2_norm[, 5:6], col = result$cluster)
```

```{r}
# Product Related, vs Product Related Duration

plot(df2_norm[, 7:8], col = result$cluster)
```

## 9 Challenging the solution

### Hierachical clustering

```{r}
# We use R function hclust() 
# For hierarchical clustering
# First we use the dist() to compute the Euclidean distance btwn obs
# d will be the first argument in the hclust() dissimilairty matrix
# 

d <- dist(df2_norm, method = "euclidean")

# We then apply hierarchical clustering using the Ward's method

res.hc <- hclust(d, method = "ward.D2")

# Lastly we plot the obtained dendrogram
#--

plot(res.hc, cex = 0.6, hang = -1)
```


